### 2016-01-23 featureImagesTable.tex
Initial commit.

\begin{table}[!htb]
  \centering
  \begin{tabular*}{0.65\textwidth}{@{\extracolsep{\fill}} ll}
    \multicolumn{1}{c}{\textbf{Feature type}} & \multicolumn{1}{c}{\textbf{Image source}} \\
    \toprule
    \midrule
    \multicolumn{2}{c}{\textbf{Intensities}} \\
    \midrule
    normalized/preprocessed & FLAIR, T1, and T2 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Symmetric template}} \\
    \midrule
    template difference & FLAIR, T1, and T2 \\
    contralateral difference & FLAIR, T1, and T2 \\
    location indices & FLAIR, T1, and T2 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Segmentation probabilities}} \\
    \midrule
    $Pr$(cerebrospinal fluid) & T1 \\
    $Pr$(gray matter) & T1 \\
    $Pr$(white matter) & T1 \\
    $Pr$(deep bray matter) & T1 \\
    $Pr$(brain stem) & T1 \\
    $Pr$(cerebellum) & T1 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Distance maps}} \\
    \midrule
    cerebrospinal fluid & T1 brain segmentation \\
    gray matter & T1 brain segmentation \\
    deep gray matter & T1 brain segmentation \\
    whole brain & T1 brain segmentation \\
    \midrule
    \multicolumn{2}{c}{\textbf{Neighborhood statistics}} \\
    \midrule
    mean & FLAIR, T1, and T2 \\
    standard deviation & FLAIR, T1, and T2 \\
    skewness & FLAIR, T1, and T2 \\
    \midrule
    \bottomrule
  \end{tabular*}
 \label{table:indices}
 \caption{List of feature images used for Stage 1 of the proposed white matter
          hyperintensity segmentation framework.}
\end{table}