### 2016-01-29 featureImagesTable.tex
ENH:  More.

\begin{table}[!htb]
  \centering
  \begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}} lll}
    \multicolumn{1}{c}{\textbf{Feature type}} & \multicolumn{1}{c}{\textbf{Image source}} & \multicolumn{1}{c}{\textbf{Name}} \\
    \toprule
    \midrule
    \multicolumn{3}{c}{\textbf{Intensities}} \\
    \midrule
    normalized/preprocessed & FLAIR, T1, and T2 & \texttt{Preprocessed} \\
    \midrule
    \multicolumn{3}{c}{\textbf{Symmetric template}} \\
    \midrule
    template difference & FLAIR, T1, and T2 & \texttt{SymmetricTemplateDifference}\\
    contralateral difference & FLAIR, T1, and T2 & \texttt{ContralateralDifference}\\
    template location indices & FLAIR, T1, and T2 & \texttt{TemplateIndicesWarped}\\
    \midrule
    \multicolumn{3}{c}{\textbf{Segmentation probabilities}} \\
    \midrule
    $Pr$(cerebrospinal fluid) & T1 & \texttt{BrainSegmentationPosteriors1}\\
    $Pr$(gray matter) & T1 & \texttt{BrainSegmentationPosteriors2}\\
    $Pr$(white matter) & T1 & \texttt{BrainSegmentationPosteriors3}\\
    $Pr$(deep gray matter) & T1 & \texttt{BrainSegmentationPosteriors4}\\
    $Pr$(brain stem) & T1 & \texttt{BrainSegmentationPosteriors5}\\
    $Pr$(cerebellum) & T1 & \texttt{BrainSegmentationPosteriors6}\\
    \midrule
    \multicolumn{3}{c}{\textbf{Distance maps}} \\
    \midrule
    cerebrospinal fluid & T1 brain segmentation & \texttt{NormalizedCsfDistance}\\
    gray matter & T1 brain segmentation & \texttt{NormalizedGmDistance}\\
    deep gray matter & T1 brain segmentation & \texttt{NormalizedDeepGmDistance}\\
    whole brain & T1 brain segmentation & \texttt{NormalizedMaskDistance}\\
    \midrule
    \multicolumn{3}{c}{\textbf{Neighborhood statistics}} \\
    \midrule
    mean & FLAIR, T1, and T2 & \texttt{NeighborhoodMean}\\
    standard deviation & FLAIR, T1, and T2  & \texttt{NeighborhoodSigma}\\
    skewness & FLAIR, T1, and T2  & \texttt{NeighborhoodSkewness}\\
    \midrule
    \bottomrule
  \end{tabular*}
 \label{table:indices}
 \caption{List of feature images used for Stage 1 of the proposed white matter
          hyperintensity segmentation framework.  These feature images encompass
          salient location and intensity information used for voxelwise identification
          of white matter hyperintensities in TBI.  The column on the right provides
          the names used in the software developed for this methodology and
          facilitates discussion of the results (cf Figure 2).
          }
\end{table}