### 2016-02-11 stitched.Rmd

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.5}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
   - \definechangesauthor[name={James Stone}, color=magenta]{js}
   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
bibliography: references.bib
csl: national-science-foundation-grant-proposals.csl
fontsize: 11pt
mainfont: Georgia
geometry: margin=1.0in
---

<!--
   - \setlength{\parskip}{3pt}
   - \setlength{\topsep}{0pt}
   - \setlength{\partopsep}{0pt}
   - \setlength{\itemsep}{0pt}
   - \setlength{\floatsep}{0pt}
   - \setlength{\intextsep}{2pt}
   - \setlength{\abovecaptionskip}{2pt}
   - \setlength{\belowcaptionskip}{0pt}
-->

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

\pagenumbering{gobble}

# Introduction

## White matter hyperintensities in TBI

White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences. These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences. WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke. WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa]. Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa]. Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa]. Additionally, the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].

Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice. Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis. Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa]. As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state. Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.

## Random forests for WMH segmentation

Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated
extensions.

The random forests framework [@breiman2001] is a popular machine learning technique that has
demonstrated significant utitility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., human gait detection [@viola2005]).
In the context of neuropathology, random forest-based paradigms have been employed in the
delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions.  In response to the lack of objective comparisons between
segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI
conference.

Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
that are built from training data.  Once constructed, data to be classified is "pushed"
through each decision tree resulting in a single classification "vote" per tree.  These
votes are then used for regression or classification of the data.  Although decision
trees had been extensively studied previously, the success of employing collections of
such weak learners for boosting machine learning performance
(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
data (i.e., "bagging") resulting in the current random forest technique.

In this work, we develop a concatenated random forest framework with a tailored contextual
feature image set (both spatial and intensity-based)
for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
the entire framework is provided publicly through the well-known open-source
ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the research of this work is
the public availability of the imaging data thus permitting full reproducibility
of the results reported and discussed.

[^i1]: https://github.com/stnava/ANTs
[^i2]: https://github.com/stnava/ANTsR

# Materials and Methods

## Imaging

MR images utilized for this initial report were acquired from 24 individuals scanned at a single site involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}). Briefly, participants were Operation Enduring Freedom/ Operation Iraqi Freedom /Operation New dawn (OEF/OIF/OND) era service members and veterans between the ages of 18--60 who have experienced combat situations and are between X and X (period of time) from deployment. \added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}

Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).  3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.  3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 x 256 mm FOV.  In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.

## Quantitative analysis

Crucial to these supervised segmentation approaches are the creation and selection of
"features" as input in conjunction with
\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
{expertly identified structures of interest}{the ground-truth} for model construction.  For the targeted application in this work (i.e., WMHs),
regression/classification are performed at the voxelwise level.  In other words, each voxel
within the region of interest is sent through the ensemble of decision trees and receives
a set of classification votes from each tree thus permitting a regression or classification
solution.  Since this procedure is performed at the voxelwise level, intensity information
alone is insufficient for good segmentation performance due to the lack of spatial context.
For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
periventricular caps in normal subjects which often confounds automated lesion detection
algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
even though machine learning and pattern recognition techniques are extremely powerful and
have significant potential, just as crucial to outcome is the creative construction and
deployment of salient feature images which we detail below.

### Feature images for WMH segmentation

\input{featureImagesTable.tex}

Supervised methodologies are uniquely characterized, in part, by the feature images that
are used to identify the regions of interest.  In Table 1, we provide a list and basic
categorization of the feature images used for the initial (i.e., Stage 1---more on the use
of multiple random forest stages below) segmentation of the white matter hyperintensities.
In addition Figure 1 provides a representation of a set of feature images for a single
subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
with seven labels:

* cerebrospinal fluid (label 1),
* gray matter (label 2),
* white matter (label 3),
* deep gray matter (label 4),
* brain stem (label 5),
* cerebellum (label 6), and
* white matter hyperintensities (label 7).

![Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are \added[id=am,remark=NT: no need for 12 dof for
intra-subject normalization]{rigidly} pre-aligned
[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.](Figures/featureImages.png)

As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
acquisitions.  The FLAIR and T2 images are rigidly registered to the T1 image using
the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
range $[0,1]$.  Although we could have used an alternative intensity standardization
algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results.

![Sample FLAIR acquisition image slices showing both manual (top) and random forest
(RDF) segmentations (bottom) obtained during the leave-one-out evaluation.  Manual
segmentations were performed by one of the authors and provided the ground
truth WMH labels for training the random forest models.](Figures/montage.png)

The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation (i.e., the pathology is discounted which is
a fairly safe assumption for the T1 image) protocols as part of the
cortical thickness pipeline described in [@Tustison:2014ab].
\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
brain has been extracted, we apply a Bayesian-based segmentation algorithm
with a template-based prior probability strategy to segment the
parenchymal tissue types.  The result is} a mask delineating the brain parenchyma and probabilistic estimates of the
cerebrospinal fluid (csf), gray matter, white matter, deep gray matter, brain stem, and
cerebellum.  These provide the expertly annotated labels for the first six tissue labels
given above.  The white matter hyperintensities were manually identified by one of the authors
(J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
the ANTs Atropos tool [@Avants:2011aa] and
multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
MMRR data set  [@landman2011] (cf Figure 3).

![Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for specific intensity-based
features.](Figures/MMRR.png)

To model the intensity information the first set of images simply includes the
preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.  We also calculate a set
of neighborhood statistics (mean, standard deviation, and skewness) feature images using
a \added[id=am]{Manhattan} radius of one voxel. For each of the normalized
images, we calculate the difference in intensities with the corresponding warped template
component.  Previous success in the international brain tumor segmentation
competition [@Menze:2015aa] was based on an important set of intensity features that were created
from multi-modal templates mentioned previously [@Tustison:2015aa].  We employ the
same strategy here.  For example, the template
difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:

$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$

where $\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$ is the transform
which maps from the individual subject space to the
template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
to take advantage of the \added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature.  For the FLAIR component, this contralateral
difference image is calculated from

$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$

where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
the symmetric template.

The segmentation probability images described above are used as feature images to provide a
spatial context for the random forest model prediction step.  Additional spatial contextual
feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
deep gray matter images.  These latter images are intended to help distinguish white matter
hyperintensities from false positives induced by the partial voluming at the gray/white
matter interface.  A third set of images are based on the voxel location within the
space of the template.  The T1 image of the subject is registered to the T1 template
component using a B-spline variant [@Tustison:2013ac] of the well-known ANTs Symmetric
Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
derived as part of the registration process, we can warp the voxel index locations
back to the space of the individual subject.    Note that this is similar in motivation to the work
of [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.

### Stacked (concatenated) random forests for improved segmentation performance

In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
concatenated supervised approach, whereby the prediction output from the first random forest
model serves as partial input for a second random forest model, can significantly improve
segmentation performance.  We do the same thing for the work described here \added[id=lw]{where we employ two stacked random forests (or two "stages")}.  The Stage 1
feature images of the training data (as described previously) are used to construct the
Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)} via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
tissue priors \added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)} for a second application of the Atropos maximum aposteriori algorithm with an
additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
the second stage we use all three aligned preprocessed images for a multivariate
segmentation.  The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.

### Code and data availability

As pointed out in a recent comprehensive multiple sclerosis lesion segmentation review
[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
is quite extensive, there were only four publicly available segmentation algorithms
at the time of writing of which none are based on supervised learning.  As we did for
our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
use the \texttt{randomForest} package [@liaw2002] using the default settings with
2000 trees per model and 500 randomly selected samples per label per image.
\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}

In addition, similar to our previous offering [^1], we plan on creating
a self-encapsulated example to showcase the proposed methodology.  The fact that
the data will also be made available through the FITBIR repository along
with the manual labelings will facilitate reproducibility on the part of the reader
as well as any interest in extending the proposed framework to other data sets.

[^1]: https://github.com/ntustison/ANTsAndArboles

### Evaluation protocol overview

In order to evaluate the protocol described, we performed a leave-one-out evaluation using
the data acquired from the 24 subjects described above.  Initial processing included the creation
of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the
manual white matter hyperintensity tracings were combined to provide the truth labels
for the training data.  The "truth" labels are the seven anatomical regions given above.

The leave-one-out procedure is as follows:

* Create Stage 1 feature images for all 24 subjects.
* For each of the 24 subjects:
    + sequester the current subject and corresponding feature images.
    + construct the Stage 1 random forest model from the remaining 23 subjects.
    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
    + the previous step produces the Stage 1 voting maps for all seven labels.
    + for each of the 23 subjects, perform a Bayesian-based segmentation with an MRF spatial
      prior using the seven voting maps are used as additional tissue priors.
    + construct the Stage 2 random forest model from all the Stage 1 feature images,
      seven voting maps, and seven posterior probability maps from the previous step.
    + send the sequestered subject through the random forest models for both stages.
    + compare the final results with the manually-defined white matter hyperintensity
      regions.

# Results

## Ranking feature importance

After performing the leave-one-out evaluation described at the end of the previous section,
we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
during the out-of-bag phase of the random forest model construction and quantifies the decrease
in prediction accuracy from omitting the specified feature.  In other words, this quantity
helps determine the importance of a particular feature and, although we save such
efforts for future work, this information provides us with guidance for future feature
pruning and/or additions.

![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
forest models for Stage 1 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
\added[id=nt]{Features are ranked in descending order of importance.}
The horizontal error bars provide the $95^{th}$ percentile
\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
stability of the feature importance across the leave-one-out models.
\added[id=nt]{At this initial stage only 31 features images are used.}](Figures/averageLeaveOneOutStage1.png)

![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
forest models for Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
\added[id=nt]{Features are ranked in descending order of importance.}
The horizontal error bars provide the $95^{th}$ percentile
\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
stability of the feature importance across the leave-one-out models.
\added[id=nt]{We augment the 31 feature images from the first stage by adding an additional
7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
segmentation for a total of 45 images for the second stage.}](Figures/averageLeaveOneOutStage2.png)

The resulting rankings for both Stages are given in Figures 4 and 5 where the values for the
separate stages are averaged over the entire corresponding model set.  In addition, we
track the variance for each feature over all models to illustrate the stability of
the chosen features during the evaluation.  This latter information is illustrated as
horizontal errors bars providing the $95^{th}$ percentile \deleted[id=bt]{(i.e., $1.96 \times \sigma$)}.
Note that the reader can cross reference Table 1 for identifying corresponding feature types
and names.

One can also use these measurements as a type of sanity check.  For example, from the Stage
1  plot, one can see that the \texttt{MeanDecreaseAccuracy} values for the location indices in the
anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are greater than
those for either the inferior-superior (i.e., \texttt{TemplateIndicesWarped0}) or the
left-right (i.e., \texttt{TemplateIndicesWarped0}) directions in the space of the symmetric
template.  \deleted[id=bt]{This is intuitive since, as discussed previously, manifestation of TBI white
matter hyperintensities can often be confused with higher intensities at the
periventricular caps in normal subjects [@Neema:2009aa]} \deleted[id=am]{whereas there does not seem to
be contralateral bias in manifestation of white matter hyperintensities in TBI.}

Additionally, it is interesting to note some of the other top performing features for Stage 1.
The contralateral difference FLAIR image is  highly discriminative over the set of
evaluation random forest models \added[id=bt]{(see Figure 6)}.  This accords with the known clinical relevance of
FLAIR images for identifying white matter hyperintensities and the fact that such
pathology does not \added[id=nt]{typically} manifest symmetrically in both hemispheres.  Interestingly, the
posterior maps for the deep gray matter are extremely important for accurate white matter
hyperintensity segmentation.  Perhaps the spatial specification of deep
gray matter aids in the removal of false positives.  Inspection of the bottom of the
plots demonstrates the lack of discriminating features associated with the
T1 image which is also well-known in the clinical literature.

![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
in (b).  (c) Contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesions.png)

As described earlier, for Stage 2, we used the output random forest voting maps from Stage
1 as both features themselves and as priors for input to a Bayesian-based segmentation with
an additional MRF spatial prior.  In Figure 5, the voting maps are labeled as
"\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
feature probability maps are labeled as "\texttt{RFBrainSegmentationPosteriors}".
The Stage 2 feature importance plot follows similar
trends as that for Stage 1 with the T1 images not contributing much to the identification
of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
alone it is intuitive that such features would be important.

## White matter hyperintensity segmentation evaluation

In Figure 7 are the segmentation comparisons derived from manual segmentations of the same
data.  Despite the large variability characteristic with manual labelings in related fields
[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
current clinical practices and the methodology proposed herein is readily adapted to refinements
in training data.  On the left of Figure 7 are the improvement in Dice values [@tustison2009],
\added[id=nt]{i.e.,}

$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$

\added[id=nt]{over all white matter hyperintensities when comparing the segmentations between the two stages
where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
represents the intersection between the manual/automated lesion pair.}  Performing
the second round of supervised learning improves these Dice values.  One can also note from the
right side of Figure 6 that the total lesion load volume illustrates a few subjects that are
severe outliers in terms of the number of false positives.  The second round helps to
correct this issue.

![Comparison with manual delineation of white matter hyperintensities.  On the left are the
calculate Dice values over all white matter hyperintensities.  Note the improvement
in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
(Right) Similar results can be seen by comparing the total lesion load volume between manual
and automated detection strategies.  Although some outliers are found after the Stage 2
processing in a couple subjects, the number of outliers caused by false positives is
significantly with the second stage processing.](Figures/llvAndDice.png)

# Discussion

<!--
Nick could you add some detail here on how this technique compares to other automated or semi automated approaches for WMH detection?
-->

<!--

* Evaluation
    * Proper context
        * We only have manual labelings from a single radiologist (
        * TBI white matter hyperintensities are more difficult to segment than MS lesions
            * individual lesions are smaller
            * lesion load is typically smaller
            * not prone to enhancement
            * less distinct from surrounding tissue ?
    * Despite the doubts regarding the efficacy of a typical segmentation validation,
      our contribution is useful because
         * the total segmentation framework is publicly available within the ANTs/ANTsR
           framework
         * although only a single site is analyzed, the feature images are site-agnostic
              * we plan to apply the current RF models directly to the other site data
              * we can then build the training database further by enlisting other experts
         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
         * The data (including manual tracings) will be made available as part of the FITBIR
           effort.

-->

The current communications describes a supervised statistical learning methodology for identifying WHMs within multimodal MR brain imaging. This effort utilized information acquired from the manual segmentation of WMHs from FLAIR images to help build two-stage ensembles of decision trees for the automated identification of these lesions. Although
only a single expert was used to produce the manual labelings, our intent is to further
refine the proposed paradigm by crowdsourcing with feedback from other experts who interact with both the data and methodology.  Also, we recognize that only a single site
was used for evaluating the proposed framework.  However, we are currently processing other site data with the models developed for this work and the results look promising
since the developed features are site-agnostic.

As far as we know, this is the first report utilizing a novel random forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs tend to be more difficult to segment than MS lesions as the former tend to be smaller with an overall smaller lesion load.  Also, enhancement protocols with the former tend to be less successful than with the latter.  As mentioned previously, the work in MS lesion segmentation is extensive
with a handful of techniques being publicly available.  Our framework is also available as open-source as part of well-known neuroimaging tools which easily allows for additions/extensions but is also, as far as we know, the first random forest-based technique available for such application.

Two major meta-analyses of WMHs have been published covering the periods of 1966-2009 and 2010-2015, respectively. The earlier meta-analysis covered 53 longitudinal studies that included samples of high-risk populations, i.e., patients selected for a specific disease or condition such as hypertension, whereas other studies recruited samples of the general population. Longitudinal studies of samples representative of the general population are more relevant to the focus of the present paper. Debette & Markus [@Debette:2010aa] found that the presence of WMHs was related to subsequent cognitive decline, a higher risk of developing dementia, stroke, and of mortality. Lesion volume at baseline was also predictive of cognitive decline. Limitations of this meta-analysis include heterogeneity in the method of measuring WMHs; some studies used automated volumetric measurement, whereas others used a visual rating scale. The studies analyzed by Debette & Markus were limited to the occurrence of one of the aforementioned conditions which they analyzed by hazard ratios.

A recent meta-analysis by Kloppenborg et al. [@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age. These studies also differed in the metric used to measure the WMHs, including volume, % of total intracranial volume, and a visual rating score. The effect size for the association with cognitive deficit in these cross-sectional studies did not differ significantly across various cognitive domains or the method of measuring lesion volume.  Among eight longitudinal studies analyzed by Kloppenborg et al that included a follow-up MRI and also controlled for age, the effect size for the association of progression in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09). This association was stronger for attention and executive function than for memory and processing speed. Although baseline WMHs were predictive of cognitive deficit at follow-up in the seven studies which did not repeat MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in the longitudinal studies that calculated progression in WMHs. In summary, progression of WMHs seen on repeat MRI has a stronger relation to cognitive deficit than concurrent imaging findings. These meta-analyses support the rationale for repeating an MRI in patients younger than 50 years whose initial scan shows WMHs.
Despite the above-described associations between WMHs, cognitive decline, increased risk of developing dementia, and mortality, these lesions receive little attention in current clinical workflows. When reported in a standard neuroradiologist interpretation, they are typically handled as incidental findings and are assigned little clinical significance. This likely reflects the impracticality of performing a detailed assessment of number, volume, and distribution within a qualitative neuroradiologist interpretation as well as the lack of correlative information on how the presence and distribution of these lesions may inform a diagnosis and prognosis in the appropriate clinical setting. To date, automated or semi-automated tools for the detection of WMHs have lacked the specificity and efficiency for the mining of large-scale datasets to generate highly granular data on whether these lesions possess any true diagnostic or prognostic value in the setting of a specific disease process. The present communication describes a supervised statistical learning tool that is appropriate for the application to such large-scale datasets.

The currently described tool is just one example of how “deep learning” algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state. It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations. Deep learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features. Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data. Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient. These approaches may be applied to the identification of macroscopically visible lesions as described in the current communication. Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments. At present, these sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient. Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases. Acquisition of these data is resource intensive and requires a multi-center approach with harmonized scanner between sites and correlative non-imaging clinical data. Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through the Federal Interagency TBI repository (FITBIR). In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.

\clearpage

\listofchanges

\clearpage

# References