### 2016-02-11 intro.Rmd

<!--
A couple notes in response to comments:

[David Tate, Comment 15]:  Stupid question, but why not show the same subject and slice for both for a better comparison?

NT:  I see what he is saying and probably agree.

[David Tate, Comment 28]:  Maybe this could go in a Figure as a flow chart??

NT:  Flow charts are great and I included a couple as Figure 5 in my tumor paper.  However, they take a long time to make correctly and I do not want to commit to producing one unless it is absolutely necessary.

[Andy Mayer, Comment 30]: It will be difficult to see this font for labeling on Y axis at current size/resolution. You may also consider alerting your reader directly that the magnitudes of the scales are very different for stage 1 versus stage 2 in the legend or scale them to be similar.

NT:  Comparing these two plots in Figure 4 in terms of their absolute MeanDecreaseAccuracy
is meaningless so coordinating the scales is not important.

[Andy Mayer, Comment 34]:  A reviewer is likely to ask for some sort of reliability rating on the labeling of the lesions.

NT:  I agree with this but unless somebody else is willing to manually segment the WMHs,
then there is not really much we can do other than point to other papers which have looked
at reliability.

[David Tate, Comment 37]:  Seems like we are missing a picture.  That is the football player in me, but it seems like we need a picture that shows the manual compared to the automated at each stage if possible.

NT:  We could do this but I do not think that such images in publications are that meaningful as they typically result from heavy selection bias.  I did do something similar
for my brain segmentation paper where I showed multiple slices (skipped every 10, or something like that) for each subject in the challenge cohort but tumors are obviously much bigger and can be displayed in this manner.

[David Tate, Comment 39]:  I would would move these comments down to a limitation like paragraph, otherwise it minimizes the cool findings.

NT:  I have no opinion on this.  Do what you think is best.

[Andy Mayer, Comment 40]:  I do not know if you can include any of this data (even one subject). It would be great to demonstrate that your algorithm worked on data collected on another scanner since the intensities are likely to be very different.

NT:  As I mentioned to you previously, I think this is a great idea and I would be happy to include it but how do we showcase this unless somebody is willing to annotate data from these other sites.  We could include a picture of a single subject but how persuasive that is will be reviewer-dependent.  Personally, I would not be persuaded as the showing the results of a single subject is very prone to selection bias.

-->

# Introduction

## White matter hyperintensities in TBI

White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences. These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences. WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke. WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa]. Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa]. Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa]. Additionally, the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].

Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice. Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis. Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa]. As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state. Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.

## Random forests for WMH segmentation

Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated
extensions.

The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant \replaced[id=dt]{utility}{utitility} for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., human gait detection [@viola2005]).
In the context of neuropathology, random forest-based paradigms have been employed in the
delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions \replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}[@Menze:2015aa].

Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
that are built from training data.  Once constructed, data to be classified is "pushed"
through each decision tree resulting in a single classification "vote" per tree.  These
votes are then used for regression or classification of the data.  Although decision
trees had been extensively studied \deleted[id=dt]{previously}, the success of employing collections of
such weak learners for boosting machine learning performance
(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
Finally, Breiman [@breiman2001] improved accuracy by random sampling of \deleted[id=dt]{the} training
data (i.e., "bagging") resulting in the current random forest technique \added[id=dt]{applied here}.

In this work, we develop a concatenated random forest framework with a tailored contextual
feature image set (both spatial and intensity-based)
for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in \replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.  Additionally,
the entire framework is provided publicly through the well-known open-source
\added[id=lw]{Advanced Normalization Tools} (ANTs)[^i1] and ANTsR[^i2] toolkits.
\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
the public availability of the imaging data thus permitting full reproducibility
of the results reported and discussed.}

[^i1]: https://github.com/stnava/ANTs
[^i2]: https://github.com/stnava/ANTsR