### 2016-02-11 methods.Rmd

# Materials and Methods

## Imaging

MR images utilized for this initial report were acquired from 24 individuals scanned at a single site involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}). Briefly, participants were Operation Enduring Freedom/ Operation Iraqi Freedom /Operation New dawn (OEF/OIF/OND) era service members and veterans between the ages of 18--60 who have experienced combat situations and are between X and X (period of time) from deployment. \added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}

Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).  3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.  3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 x 256 mm FOV.  In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.

## Quantitative analysis

Crucial to these supervised segmentation approaches are the creation and selection of
"features" as input in conjunction with
\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
{expertly identified structures of interest}{the ground-truth} for model construction.  For the targeted application in this work (i.e., WMHs),
regression/classification are performed at the voxelwise level.  In other words, each voxel
within the region of interest is sent through the ensemble of decision trees and receives
a set of classification votes from each tree thus permitting a regression or classification
solution.  Since this procedure is performed at the voxelwise level, intensity information
alone is insufficient for good segmentation performance due to the lack of spatial context.
For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
periventricular caps in normal subjects which often confounds automated lesion detection
algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
even though machine learning and pattern recognition techniques are extremely powerful and
have significant potential, just as crucial to outcome is the creative construction and
deployment of salient feature images which we detail below.

### Feature images for WMH segmentation

\input{featureImagesTable.tex}

Supervised methodologies are uniquely characterized, in part, by the feature images that
are used to identify the regions of interest.  In Table 1, we provide a list and basic
categorization of the feature images used for the initial (i.e., Stage 1---more on the use
of multiple random forest stages below) segmentation of the WMHs.
In addition Figure 1 provides a representation of a set of feature images for a single
subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
with seven labels:

* cerebrospinal fluid (label 1),
* gray matter (label 2),
* white matter (label 3),
* deep gray matter (label 4),
* brain stem (label 5),
* cerebellum (label 6), and
* white matter hyperintensities (label 7).

![Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are \added[id=am,remark=NT: no need for 12 dof for
intra-subject normalization]{rigidly} pre-aligned
[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.](Figures/featureImages.png)

As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
acquisitions.  The FLAIR and T2 images are rigidly registered to the T1 image using
the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
range $[0,1]$.  Although we could have used an alternative intensity standardization
algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar \added[id=dt]{to previous work}[@Tustison:2015aa].

![Sample FLAIR acquisition image slices showing both manual (top) and random forest
(RDF) segmentations (bottom) obtained during the leave-one-out evaluation.  Manual
segmentations were performed by one of the authors and provided the ground
truth WMH labels for training the random forest models.](Figures/montage.png)

The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation (i.e., the pathology is discounted which is
a fairly safe assumption for the T1 image) protocols as part of the
cortical thickness pipeline described in [@Tustison:2014ab].
\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
brain has been extracted, we apply a Bayesian-based segmentation algorithm
with a template-based prior probability strategy to segment the
parenchymal tissue types.  The result is} a mask delineating the brain parenchyma and probabilistic estimates of the
\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white matter, deep gray matter, brain stem, and
cerebellum.  These provide the expertly annotated labels for the first six tissue labels
given above.  The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually identified by one of the authors
(J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
the ANTs Atropos tool [@Avants:2011aa] and
multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
MMRR data set  [@landman2011] (cf Figure 3).

![Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for specific intensity-based
features.](Figures/MMRR.png)

To model the intensity information the first set of images simply includes the
preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.  We also calculate a set
of neighborhood statistics (mean, standard deviation, and skewness) feature images using
a \added[id=am]{Manhattan} radius of one voxel \added[id=nt]{given the typical size
of individual WMHs}. For each of the normalized
images, we calculate the difference in intensities with the corresponding warped template
component.  Previous success in the international brain tumor segmentation
competition [@Menze:2015aa] was based on an important set of intensity features that were created
from multi-modal templates mentioned previously [@Tustison:2015aa] \added[id=nt]{and listed in Table 1}.  We employ the
same strategy here.  For example, the template
difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:

$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$

where $\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$ is the transform
which maps from the individual subject space to the
template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
to take advantage of the \added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature.  For the FLAIR component, this contralateral
difference image is calculated from

$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$

where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
the symmetric template.

The segmentation probability images described above are used as feature images to provide a
spatial context for the random forest model prediction step.  Additional spatial contextual
feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
deep gray matter images.  These latter images are intended to help distinguish white matter
hyperintensities from false positives induced by the partial voluming at the gray/white
matter interface.  A third set of images are based on the voxel location within the
space of the template.  The T1 image of the subject is registered to the T1 template
component using a B-spline variant [@Tustison:2013ac] of the well-known ANTs Symmetric
Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
derived as part of the registration process, we can warp the voxel index locations
back to the space of the individual subject \replaced[id=nt]{which motivates similar work by others}{
Note that this is similar in motivation to the work
of} [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.

### Stacked (concatenated) random forests for improved segmentation performance

In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
concatenated supervised approach, whereby the prediction output from the first random forest
model serves as partial input for a second random forest model, can significantly improve
segmentation performance.  We do the same thing for the work described here \added[id=lw]{where we employ two stacked random forests (or two "stages")}.  The Stage 1
feature images of the training data (as described previously) are used to construct the
Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)} via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
tissue priors \added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)} for a second application of the Atropos maximum aposteriori algorithm with an
additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
the second stage we use all three aligned preprocessed images for a multivariate
segmentation.  The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.

### Code and data availability

As pointed out in a recent comprehensive multiple sclerosis lesion segmentation review
[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
is quite extensive, there were only four publicly available segmentation algorithms
at the time of writing \added[id=am]{this article}.  \replaced[id=dt]{In contrast to the current work,}{of which} none are based on supervised learning.  As we did for
our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
use the \texttt{randomForest} package [@liaw2002] using the default settings with
2000 trees per model and 500 randomly selected samples per label per image.
\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}

In addition, similar to our previous offering [^1], we plan on creating
a self-encapsulated example to showcase the proposed methodology.  The fact that
the data will also be made available through the \replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR} repository along
with the manual labelings will facilitate reproducibility on the part of the reader
as well as any interest in extending the proposed framework to other data sets.

[^1]: https://github.com/ntustison/ANTsAndArboles

### Evaluation protocol overview

In order to evaluate the protocol described, we performed a leave-one-out evaluation using
the data acquired from the 24 subjects described above.  Initial processing included the creation
of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the
manual white matter hyperintensity tracings were combined to provide the truth labels
for the training data.  The "truth" labels are the seven anatomical regions given above.

The leave-one-out procedure is as follows:

* Create Stage 1 feature images for all 24 subjects.
* For each of the 24 subjects:
    + sequester the current subject and corresponding feature images.
    + construct the Stage 1 random forest model from the remaining 23 subjects.
    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
    + the previous step produces the Stage 1 voting maps for all seven labels.
    + for each of the 23 subjects, perform a Bayesian-based segmentation with an MRF spatial
      prior using the seven voting maps are used as additional tissue priors.
    + construct the Stage 2 random forest model from all the Stage 1 feature images,
      seven voting maps, and seven posterior probability maps from the previous step.
    + send the sequestered subject through the random forest models for both stages.
    + compare the final results with the manually-defined white matter hyperintensity
      regions.