### 2016-02-18 abstract.Rmd
ENH:  more.

# Abstract

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with tailored features for
segmenting WMHs in a TBI cohort. The framework is provided publicly
through the Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Lesion load and overlap evaluative comparisons are
complimented by feature rankings which showcase the utility of the
concatenated approach.  Our findings suggest supervised learning methods
may be applied to quantify WMHs on routine brain imaging. Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

\clearpage