### 2016-02-25 acknowledgementsDisclaimers.Rmd
Final submission.

## Acknowledgements

The authors wish to acknowledge all other members of
the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
and Kevin Wilson. We would also like to thank the study participants and
their family members. This project was jointly supported by the
Department of Defense (W81XWH-13-2-0095), the U.S. Department of
Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
Grant HU 0001-08-0001.

## Declaration of Interest/Disclaimer

The authors report no financial
disclosures or conflicts of interest. The views expressed here are those
of the authors and do not necessarily reflect the official policy of
position of the Department of the Navy, Department of Defense, nor the
U.S. Government. This work was prepared as a part of official duties;
Title 17 USC §105 provides that Copyright protection under this title is
not available for any work of the U.S. Government. Title 17 USC §101
defines a US Government work as a work prepared by a military service
member of employee of the US Government as part of that person’s
official duties.

\clearpage

<!-- \listofchanges

\clearpage -->

# References