### 2016-04-29 Presentation/wmh.Rmd
ENH:  Add presentation.

## WMH segmentation

![](./Figures/lesions.png)

## Feature images

$label \sim_{RF} feature_{1} + \ldots + feature_{n}$

<BR><BR>

![](../Figures/featureImages.png)

## Feature importance

![](../Figures/averageLeaveOneOut.png)

## Sample results

![](../Figures/sampleResults.png)