### 2016-04-29 Presentation/wmhBuild.Rmd

---
title: ANTs and WMHs
author: James Stone and Nick Tustison
bibliography:
logo:
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  structurebold
    highlight: tango
    incremental: false
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
    toc: no
  pdf_document:
    toc: true
    highlight: zenburn
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

##  ANTs beginnings:  image registration

<BR><BR>

![](./Figures/highresdiffeos.jpg)

##  ANTs tools

* MRI bias correction
* image denoising
* image and point set registration
* $n$-tissue segmentation
* template building
* multi-atlas label fusion
* brain extraction
* cortical thickness estimation
* ANTsR for statistics and visualization

publicly available:  https://github.com/stnava/ANTs

## Competitions

![](./Figures/evalhistory.png)

## Competition:  BRATS 2013

![](./Figures/bratsData.png)

*_only team out of 20 to make code publicly available._

## WMH segmentation

![](./Figures/lesions.png)

## Feature images

$label \sim_{RF} feature_{1} + \ldots + feature_{n}$

<BR><BR>

![](../Figures/featureImages.png)

## Feature importance

![](../Figures/averageLeaveOneOut.png)

## Sample results

![](../Figures/sampleResults.png)