### 2016-06-22 responseToReviewers.Rmd
ENH:  Respnose to reviewers.

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
bibliography: references.bib
csl: council-of-science-editors.csl
fontsize: 12pt
mainfont: Georgia
---

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

# Response to reviewers

_\textcolor{blue}{We appreciate the time spent by the editors and reviewers in
assessing our manuscript.  Please see below for a point-by-point response to the
issues raised.}_

## Reviewer 1

Thank you for the opportunity to review the paper "Supervised learning
technique for the automated identification of white matter
hyperintensities in traumatic brain injury". In this study, the authors
present a machine learning algorithm for automated segmentation of white
matter hyperintensities, a form of white matter degeneration visible in
FLAIR-MRI that is observed in a wide range of clinical syndromes, and
even in healthy controls. The method proposed by the authors relies on
training random forests to learn the relationship between voxel
intensities in a set of feature images derived from multimodal MRI (T1,
T2, FLAIR) and lesions drawn manually by an expert. Since the method uses
examples to learn what constitutes a lesion, this is correctly considered
a supervised method. The method is applied on a set of 24 patients with
clinically confirmed traumatic brain injury, and results are obtained
with a leave-one-out cross validation.

Overall, the method proposed by the authors is innovative and thoughtful,
and has great potential for bridging the gap between research findings
and clinical needs. The method uses a series of cutting edge solutions to
the segmentation problem: (1) multimodal imaging inputs, which is known
to increase segmentation accuracy, (2) random forests, which are an
excellent tool for finding multivariate nonlinear relationships, (3)
voxel neighborhood information, which provide contextual information for
each voxel, and (4) an algorithm that improves the segmentation in
stages. However, the manuscript itself, to my opinion, is not well
focused and needs significant improvement. Below are my specific comments
in support of this.

> _\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
> appreciate the detailed critique below to help us improve specific aspects of the
> presentation.}_

Major comments:

1. The biggest issue I had while reading the manuscript is the lack of a
coherent theme of what this paper is about. The authors span their exposé
from multiple sclerosis to the automobile industry, without stating why
these links to other diseases (or industries) have any relevance. It
appears in some parts that the closest disease with similar white matter
lesions is multiple sclerosis, but this is not clearly stated, nor the
two types of lesions are compared. It is, therefore, unclear why counting
how many MS lesion segmentation methods are publicly available is of any
importance. When explaining methods, the authors mention the utility of
support vector machines, or the existence of a python package called
SciPy. Yet, it is unclear why these things have any importance, unless
the authors make connections to their topic/method.

> _\textcolor{blue}{Nick}_

2. Second, the manuscript has no results. All the results section is focused
on talking about feature ranking, which by the way is explained in the
Results section instead of the Methods section. The most important
results (the prediction accuracy of automated segmentation), are not
mentioned in the manuscript but are only plotted in Figure 7. My advice
is to put the focus of the results section on accuracy, i.e., give
precise dice values and perhaps add other measures (metric displacement,
Hausdorf distance, sensitivity, specificity, etc.). Also, the gain in
accuracy from stage 1 to stage 2 cannot be guessed by looking at a plot,
the authors must provide a statistical comparison. When investigating
accuracy, the error in lesion size should be reported, both in terms of
volume and in terms of individual lesions counts (i.e. out of 10 lesions
9 were found, 1 was missed, etc.). This information is crucial to
clinicians, if the authors foresee the use of the method in a clinical
setting, as it seems from the discussion. Beside adding results to the
manuscript, it is equally important to give some results in the abstract.
The abstract must be self-sufficient in reporting all the major points of
the paper.

> _\textcolor{blue}{Nick}_

3. Third, there is no information about the manual lesion tracing procedure.
Were lesion drawn on FLAIR or were other modalities used as well? If
FLAIR was used as reference it is not surprising that FLAIR emerges as
one of the top predictive modalities. Descriptive statistics on the
lesions are also necessary: how big were the lesions, how many lesions
were found in each patient (the range 1-20 is not sufficiently
informative), where were the lesions located? This information helps
understand why, for example, brain stem or cerebellum labels were not
that important for achieving good lesion predictions.

> _\textcolor{blue}{Nick}_

4. Fourth, although the authors claim the method is available, I couldn't
find any link for the public. Note, it would be equally important if the
publication of the method online is associated with proper documentation
on how a new (naive) user can apply it.

> _\textcolor{blue}{Nick}_

Minor comments:

5. In the abstract, the authors mention the creation of tailored features
for segmenting WMHs. My understanding is that the method proposed by the
authors has no tailoring (if we think of tailoring as custom cutting to
WMH needs). There was no feature selection process all the features were
included as predictors. A similar pitfall if found on pg. 6 ln 56, where
the manuscript reads ``Crucial to these supervised segmentation approaches
are the creation and selection of features as input in conjunction with
expertly identified structures.'' What are the structures identified by
the expert in this study?

6. The methods section is very confusing. The section "Feature images for
WMH segmentation" starts with no explanation of what is Stage 1. The
logical flow would require template registration explained first, and all
feature computation performed later, and finally RF stages can be
explained with the available features. The pipeline also seem to be
repetitive, the N4 correction is mentioned in ln. 26 and then again in
ln. 43, which sounds like it was performed twice. Tissue segmentation is
mentioned in ln. 39, then it is stated is a Bayesian method (ln 45), and
then it is referred to as Atropos in the next page. It is unclear what is
³the first set of images² in pg. 9, or if there is a second set. Also,
mathematical formulas seem superfluous, particularly since it takes more
space to explain the formula than to say in plain English that
transformation matrices were applied to bring subject¹s images in
template space.

7. Random forests are explained in a way that can be understood only by
readers who already know how they work; a new reader (especially a
clinician) would not be able to understand the method. For example, data
pushed onto a tree is does not mean much, unless the reader knows what is
an identification tree.

8. The relationship between WMH and GOS-E is mentioned in pg. 2 but the
reader is not informed what is GOS-E. A similar issue is met few lines
after, where the ``outcome'' is mentioned, but there is no information what
is referred to as outcome.

> \textcolor{blue}{James}

9. In pg. 3 the authors propose that automated methods ``may allow for
identification of correlative patterns between WMH number, volume,
distribution, and disease state''. Some of these correlations have been
already identified in the literature, as the authors mention few lines
before. Moreover, the authors do not report number, volume, or
distribution of the lesions, to back up the utility of their method in
this regard.

10. In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think this
is the right verb.

> _\textcolor{blue}{We agree that the sentence in question could be read this way.
We have, therefore, clarified the sentence in question to read
Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
through each decision tree resulting in a single classification ``vote'' per tree.}_

> A similar issue is in pg 5 ln 52 where it is stated that ``feature images
> consisted of 26 subjects.'' It is not ethical to consider participants as feature
> images.

> _\textcolor{blue}{It was a grammatical mistake, not a question of
ethics.  It has been fixed: ``The feature images
were derived from MR acquisitions of 26 subjects\ldots''}_

11. It is unclear to me the meaning of the sentence: ``Since the inverse
transform is also derived as part of the registration process, we can
warp the voxel index locations back to the space of the individual
subject which motivates similar work by others [43].''

12. Why was Stage 1 segmentation based only on T1, while Stage 2 segmentation
was based on multivariate segmentation using all three modalities?

13. Similar to the feature ranking method, which is explain in Results, Dice
is also explained in Results. Also, it is unclear why finding which
direction of space is informative for the segmentation process would be a
``sanity check.''

14. The discussion starts with an ``intent to further refine the proposed
paradigm,'' and proceeds by mentioning other work currently being
performed which looks promising. The opening of the discussion with such
statements truly make the manuscript sound like an incomplete effort that
is not ready for publication.

> _\textcolor{blue}{We appreciate the positive view of our current work.
The full sentence in question is ``Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.''
In our opinion, simply mentioning specific details regarding future work
does not mean that the manuscript ``sounds like an incomplete effort''.
Rather, it is simply an admission, similar to much of what is reported
in the literature, that our research program is ongoing.}_

15. The discussion also spends several paragraphs on defining literature that
shows how WMHs are related to cognitive performance, or how WMHs can be
applied in diagnosing TBI. Yet, this study makes no effort to test these
efforts in these 24 patients. Consequently, those paragraphs sound
wishful and speculative. Given the methodological nature of this paper, a
more appropriate discussion would perhaps focus on the comparison with
other methods in terms of accuracy, similarities, differences, etc.

> _\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
to discuss.}_

## Reviewer: 2

Interesting paper which offers those without a current background in
state-of-the-art brain analysis tools. A good introduction to their capabilities.

1. However the authors do little to establish clinical utility of the
algorithm.

> _\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
> the opportunity to clarify the precise nature of our contribution which is methodological
> (vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
> (as established by previous research), we believe that the automatic identification
> of WMHs is of sufficient merit for publication.  However, our follow-up research plans
> concern the application of our work to a much larger and investigation into the clinical
> correlates of WMH load.}_

2. A graphical description of the pipeline would help most
readers to understand how this might get into the clinic.

> _\textcolor{blue}{Done.  We added the following figure as a graphical description of
> the pipeline to the manuscript:}_

![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)