### 2016-06-30 results.Rmd

# Results

## White matter hyperintensity segmentation evaluation

\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
of the previously described TBI data over the three lesion volume ranges.
These performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
size ranges over which these measures are computed are meant to illustrate the variation
in performance with lesion size.}

\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
to identify which is why the sensitivity for this range is more varied compared with the largest
set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
the volumetric estimate when comparing the manually identified lesions versus the
automatically predicted lesions.}

\textcolor{blue}{
We averaged these measures over all lesions and all subjects which resulted in the
following values:}

* \textcolor{blue}{Stage 1}

    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34

    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36

    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36

    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%

* \textcolor{blue}{Stage 1}

    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38

    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40

    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36

    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%

<!--

In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
data.  Despite the large variability characteristic with manual labelings in related fields
[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
current clinical practices and the methodology proposed herein is readily adapted to refinements
in training data.

On the left of Figure 5 are the improvement in Dice values [@tustison2009],
over all white matter hyperintensities when comparing the segmentations between the two stages
where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
represents the intersection between the manual/automated lesion pair.  Performing
the second round of supervised learning improves these Dice values.  One can also note from the
right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
severe outliers in terms of the number of false positives.  The second round helps to
correct this issue.

-->

![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)

## Ranking feature importance

After performing the leave-one-out evaluation,
we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
during the out-of-bag phase of the random forest model construction and quantifies the decrease
in prediction accuracy from omitting the specified feature.  In other words, this quantity
helps determine the importance of a particular feature and, although we save such
efforts for future work, this information provides us with guidance for future feature
pruning and/or additions.

![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 1 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
 and illustrate the
stability of the feature importance across the leave-one-out models.
At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)

![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile and illustrate the
stability of the feature importance across the leave-one-out models.
We augment the 31 feature images from the first stage by adding an additional
7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)

The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
separate stages are averaged over the entire corresponding model set.  In addition, we
track the variance for each feature over all models to illustrate the stability of
the chosen features during the evaluation.  This latter information is illustrated as
horizontal errors bars providing the $95^{th}$ percentile
Note that the reader can cross reference Table 1 for identifying corresponding feature types
and names.