### 2016-12-02 methodsX.Rmd

# Materials and Methods

## Imaging

MR images utilized for this initial report were acquired from a single
scanner involved in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study (see Walker et al., this issue). Briefly,
participants were Operation Iraqi Freedom/Operation Enduring Freedom
(OIF/OEF) era Service Members and Veterans between the ages of 18-60
years with prior combat exposure and deployment(s). The feature images
\textcolor{black}{were derived from MR acquisitions of} 26 subjects aged
39.6 $\pm$ 8.1 years (range 28--58 years). Within this cohort, 24 (92%)
were considered positive for TBI based upon the potential concussive
events (PCE) interview process described in detail in Walker et al.,
this issue. \textcolor{black}{All lesions were isolated in the white
matter of the cerebrum.  Table 1 provides a descriptive statistical
summary of the variation in lesion load across the selected cohort.}

\input{trainingData.tex}

Images were acquired on a Philips 3.0T Ingenia system with an 8-channel
SENSE head coil (Philips Medical Systems, Best, Netherlands).  3D FLAIR
sequences were acquired with a turbo spin echo inversion recovery
sequence with the following parameters:  repetition time (TR) = 4800 ms,
echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal
slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition
matrix, and 256 $\times$ 256 mm FOV.  3D T1-weighted sequences were
acquired with a fast field echo (FFE) sequence with the following
parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240;
flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice
thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.  In
addition, 3D T2-weighted images were acquired with a turbo spin echo
sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ETL:
= 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$
256 acquisition matrix, and 256 x 256 mm FOV.

\textcolor{black}{The first author (J. R. S.) performed the manual WMH
tracings for all 26 subjects.  J. R. S. is a radiologist certified by
the American Board of Radiology, with a certificate of advanced
qualification in vascular and interventional radiology, over 18 years of
research experience in TBI, and 6 years of clinical imaging experience.
All multi-modal MR dicom image slices were converted to the nifti file
format} [^1]. \textcolor{black}{All nifti image volumes for each subject
were rigidly aligned to the T1 image of that subject using the ANTs
software} [@Avants:2014aa] \textcolor{black}{.  The normalized MRI
volumes were then provided to J. R. S. who traced each lesion using the
ITK-SNAP tool} [@Yushkevich:2006aa] \textcolor{black}{ which has
multi-image overlay capabilities for visualizing all modalities in all
three canonical views.}

[^1]: http://nifti.nimh.nih.gov/nifti-1

## Quantitative analysis

\textcolor{black}{Figure 1 provides a graphical overview of the proposed
workflow.  The major components include offline generation of symmetric
multimodal templates, the creation of feature images from the training
data which are then employed for statistical prediction using a
concatenated random forest framework.  This framework involves the use
of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application, which we
denote as ``Stage1'', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as
``Stage 2''. Once these offline steps are performed, a new (i.e.,
unsegmented) subject can then be processed using the proposed pipeline.}

![\textcolor{black}{Workflow illustration for the proposed pipeline.
Processing of the multi-modal input MRI for a single subject, using the
multi-modal symmetric template, results in the generation of the feature
images.  These feature images are used as input to the Stage 1 RF model
producing the initial RF probability map estimates.  The Stage 1 voting
maps, the original feature images, and the Stage 2 RF model result in
the final voting maps which includes the WMH probability estimate.  Note
that the RF models are constructed once from a set of training data
which are processed using the same feature-construction pipeline as the
single-subject input MRI.}](Figures/wmhPipeline.png)

### Symmetric multi-modal templates

\textcolor{black}{Following} [@Tustison:2014ab] \textcolor{black}{and}
[@Tustison:2015aa], \textcolor{black}{optimally derived templates serve
for both brain tissue segmentation (for the derivation of tissue prior
probability maps) and generation of asymmetry feature images. For this
work we use the multi-modal data available from the public MMRR data
set} [@landman2011]\textcolor{black}{.  We used all 46 multi-modal
acquisitions of that study to produce a multi-modal template according
to the procedure described in} [@Avants:2010aa] \textcolor{black}{which
results in a mean (in terms of both shape and intensity) multivariate
template representing the entire cohort.  Mid-canonical slices of the
FLAIR, T1, and T2 components are illustrated in Figure 2.}

![Canonical views of the mutlivariate, bilaterally symmetric template
constructed from the MMRR data set [@landman2011] (only shown are the
FLAIR, T1, and T2 modalities--- the components relevant for this work).
Template construction is detailed in [@Tustison:2015aa].  These images
are important for specific asymmetry-based features.](Figures/MMRR.png)

### Feature images for WMH segmentation

\textcolor{black}{ Crucial to these supervised segmentation approaches
are the creation and selection of ``features'' as input (i.e., feature
images constructed from the training data) in conjunction with expertly
identified structures of interest (i.e., WMHs) for model construction.}
For the targeted application in this work, tissue classification is
performed at the voxelwise level.  In other words, each voxel within the
region of interest is sent through the ensemble of decision trees and
receives a set of classification votes from each tree thus permitting a
regression or classification solution.  Since this procedure is
performed at the voxelwise level, intensity information alone is
insufficient for good segmentation performance due to the lack of
spatial context. For example, as pointed out in [@Neema:2009aa], higher
intensities can be found at the periventricular caps in normal subjects
which often confounds automated lesion detection algorithms.  Other
potential confounds include MR signal inhomogeneity and noise.
Therefore, even though machine learning and pattern recognition
techniques are extremely powerful and have significant potential, just
as crucial to outcome is the creative construction and deployment of
salient feature images which we detail below.

\input{featureImagesTable.tex}

Supervised methodologies are uniquely characterized, in part, by the
feature images that are used to identify the regions of interest.  In
Table 2, we provide a list and basic categorization of the feature
images used for the initial (i.e., Stage 1---more on the use of multiple
random forest stages below) segmentation of the WMHs. In addition Figure
3 provides a representation of a set of feature images for a single
subject analyzed in this work.  Note that in this work we categorize the
brain parenchyma with seven labels:

* cerebrospinal fluid (label 1),
* gray matter (label 2),
* white matter (label 3),
* deep gray matter (label 4),
* brain stem (label 5),
* cerebellum (label 6), and
* white matter hyperintensities (label 7).

![Representation of Stage 1 feature images for subject 9.  The FLAIR,
T1-, and T2-weighted images are rigidly pre-aligned [@Avants:2014aa] to
the space of the FLAIR image.  The three modality images are then
preprocessed (N4 bias correction [@Tustison:2010ac] and adaptive
denoising [@Manjon:2010aa]) followed by application of standard ANTs
brain extraction and $n$-tissue segmentation protocols using the MMRR
symmetric template and corresponding priors [@Tustison:2014ab] applied
to the T1 image. The feature images are then generated for voxelwise
input to the RF model which results in the voting maps illustrated on
the right.  This gives a probabilistic classification of tissue type.
Not shown are the probability and voting images for the brain stem and
cerebellum.](Figures/featureImages.png)

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions.  The T1 and T2 images are rigidly
registered to the FLAIR image using the open-source Advanced
Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images are
then preprocessed using the denoising algorithm of [@Manjon:2010aa]
followed by N4 bias correction [@Tustison:2010ac] which are then
normalized to the intensity range $[0,1]$.  Although we could have used
an alternative intensity standardization algorithm (e.g., [@nyul2000]),
we found that a simple linear rescaling produced better results similar
to previous work[@Tustison:2015aa].

![Sample FLAIR acquisition image slices showing both manual and random
forest segmentations for both stages obtained during the leave-one-out
evaluation.  Manual segmentations were performed by one of the authors
and provided the ground truth WMH labels for training the random forest
models.](Figures/sampleResults.png)

The T1 image is then processed via the ANTs brain extraction and normal
tissue segmentation pipelines [@Tustison:2014ab]. The result is a mask
delineating the brain parenchyma and probabilistic estimates of the CSF,
gray matter, white matter, deep gray matter, brain stem, and cerebellum
[@Avants:2011aa].  These provide the expertly annotated labels for the
first six tissue labels given above.  Tissue prior probability maps for
segmentation are from multi-model optimal symmetric shape/intensity
templates [@Tustison:2015aa] created from the public MMRR data set
[@landman2011] (cf Figure 2).

Feature values include the preprocessed FLAIR, T1, and T2 image voxel
intensities. We also calculate a set of neighborhood statistics (mean,
standard deviation, and skewness) feature images using a Manhattan
radius of one voxel given the typical size of individual WMHs. For each
of the preprocessed images, we calculate the difference in intensities
with the corresponding warped template component.  Previous success in
the international brain tumor segmentation competition [@Menze:2015aa]
was based on an important set of intensity features that were created
from multi-modal templates mentioned previously [@Tustison:2015aa] and
listed in Table 2.  We employ the same strategy here.

<!-- For example, the template difference feature image for the FLAIR
image, $S_{FLAIR}$ is calculated as:

$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$

where

$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$

is the transform which maps from the individual subject space to the
template space and $T_{FLAIR}$ is the FLAIR template component. -->

To take advantage of the gross bilateral symmetry of the normal brain
(in terms of both shape and intensity), and the fact that WMHs do not
generally manifest symmetrically across hemispheres, we use the
symmetric templates to compute the contralateral intensity differences
as an additional intensity feature.

<!-- For the FLAIR component, this contralateral difference image is
calculated from

$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$

where $\phi_R$ denotes a horizontal reflection perpendicular to the
mid-sagittal plane of the symmetric template. -->

The segmentation probability images described above are used as feature
images to provide a spatial context for the random forest model
prediction step.  Additional spatial contextual feature images include
the distance maps [@maurer2003] based on the csf, gray matter, and deep
gray matter images.  These latter images are intended to help
distinguish white matter hyperintensities from false positives induced
by the partial voluming at the gray/white matter interface.  A third set
of images are based on the voxel location within the space of the
template. \textcolor{black}{Similar feature images were used in}
[@Anbeek:2004aa] \textcolor{black}{although, unlike the proposed
framework, this previous work lacks normalization to the standard
coordinate system provided by the template to dramatically improve
spatial specificity across all subjects.  To generate these images, the
T1 image of each subject is registered to the T1 template component
using a B-spline variant} [@Tustison:2013ac] \textcolor{black}{of the
well-known ANTs Symmetric Normalization (SyN) algorithm}
[@Avants:2011ab]. \textcolor{black}{Using the derived transforms, the
template coordinate images are warped back to the space of the
individual subject.}

### Stacked (concatenated) random forests for improved segmentation performance

In previous brain tumor segmentation work [@Tustison:2015aa], it was
demonstrated that a concatenated supervised approach, whereby the
prediction output from the first random forest model serves as partial
input for a second random forest model, can significantly improve
segmentation performance.  We do the same thing for the work described
here where we employ two stacked random forests (or two "stages").  The
Stage 1 feature images of the training data (as described previously)
are used to construct the Stage 1 model.  The training data Stage 1
features are then used to produce the voxelwise "voting maps" (i.e., the
classification count of each decision tree for each tissue label) via
the Stage 1 random forest model.  All the Stage 1 features plus the
Stage 1 voting maps are used as input to the Stage 2 model.  In
addition, we use the Stage 1 voting maps as tissue priors (i.e.,
probabilistic estimates of the tissue spatial locations) for a second
application of the $6$-tissue segmentation algorithm with an additional
Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
\textcolor{black}{In order to maximize the spatial information for the
$n$-tissue segmentation process following the voxelwise RF
classification of Stage 1, we use all three aligned preprocessed images
for multivariate segmentation during the second stage.} The resulting
seven posterior probability images constitute a third additional feature
image set for Stage 2.

### Implementation

As pointed out in a recent comprehensive lesion segmentation review
[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in
the literature is quite extensive, there were only four publicly
available segmentation algorithms at the time of writing this article.
In contrast to the current work, none are based on supervised learning.
As we did for our brain tumor segmentation algorithm [@Tustison:2015aa],
all of the code described in this work is publicly available through the
open-source ANTs/ANTsR toolkits.  Through ANTsR (an add-on toolkit
which, in part, bridges ANTs and the R statistical project) we use the
_randomForest_ package [@liaw2002] using the default settings with 2000
trees per model and 500 randomly selected samples per label per image.
Note that we saw little variation in performance when these parameters
were changed (i.e. up to 1000 random samples and as little as 1000
trees) which is consistent with our previous experience.

In addition, similar to our previous offering [^2], we plan on creating
a self-encapsulated example to showcase the proposed methodology
\textcolor{black}{which will also be available on github.}[^3]  The fact
that the data will also be made available through the Federal
Interagency Traumatic Brain Injury Research (FITBIR) repository along
with the manual labelings will facilitate reproducibility on the part of
the reader as well as any interest in extending the proposed framework
to other data sets.

[^2]: https://github.com/ntustison/ANTsAndArboles

[^3]: https://github.com/ntustison/WatchMeHyperventilate

### Evaluation protocol overview

In order to evaluate the protocol described, we performed a
leave-one-out evaluation using the data acquired from the 24 subjects
described above.  Initial processing included the creation of all Stage
1 feature images for all subjects.  The initial brain segmentation of
each T1 image and the manual white matter hyperintensity tracings were
combined to provide the truth labels for the training data.  The
``truth'' labels are the seven anatomical regions given above.

The leave-one-out procedure is as follows:

* Create Stage 1 feature images for all 24 subjects.
* For each of the 24 subjects:
    + sequester the current subject and corresponding feature images.
    + construct the Stage 1 random forest model from the remaining 23 subjects.
    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
    + the previous step produces the Stage 1 voting maps for all seven labels.
    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
      MRF spatial
      prior using the seven voting maps as additional tissue priors.
    + construct the Stage 2 random forest model from all the Stage 1 feature images,
      seven voting maps, and seven posterior probability maps from the previous step.
    + send the sequestered subject through the random forest models for both stages.
    + compare the final results with the manually-defined white matter hyperintensity
      regions.

\textcolor{black}{Several measures have been employed in the literature
for evaluating automated white matter lesion segmentation involving such
quantities as lesion load, voxel-based overlap measures (such as the
Dice similarity coefficient), and lesion-based
measures.}[@Garcia-Lorenzo:2013aa]\textcolor{black}{.  For this work,
due to the relatively small-size distribution} \textcolor{black}{of the
lesion load in our data cohort (see Table 1), we used four lesion-based
measures:  sensitivity, positive predictive value (i.e., precision),
$F_1$ score, and relative volume difference.  The first three quantities
are based on the number of false positives \textcolor{blue}{($FP$), true negatives ($TN$)},
false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It
should be noted that the number of true negatives is not readily
incorporated into measures of accuracy as the quantity of ``true
negatives'' (i.e., correctly identified normal brain tissue) would
severely skew accuracy assessments. The $F_1$ score is an assessment of
accuracy which takes into account both the sensitivity,}

$$ sensitivity =  \frac{ TP }{TP + FN}, $$

\textcolor{black}{and the positive predictive value (PPV),}

$$ PPV =  \frac{ TP }{TP + FP}, $$

\textcolor{black}{such that $F_1$ is given by}

$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$

\textcolor{black}{The relative volume difference is calculated for each
of the true positive lesions using the manual and predicted lesion
volumes:}

$$ Relative\,\,volume\,\,difference = \frac{V_{manual} -
V_{predicted}}{V_{manual}}. $$

\textcolor{black}{To illustrate how the performance of our framework
varies with lesion size, we calculated the above measures based on
evenly split quantiles of the manual estimates of lesion volumes into 3
groups.  These three size ranges (in terms of the number of voxels)
are:}

* \textcolor{black}{Quantile 1: $[1-12)$,}

* \textcolor{black}{Quantile 2: $[12-28)$, and}

* \textcolor{black}{Quantile 3: $[28-551]$}.

\textcolor{black}{These quantiles are used to showcase the performance
(cf Figure 5).}