### 2017-08-06 Presentation/wmh.Rmd
ENH:  Add slides for Lisa.

## WMH segmentation

\begin{center}
\includegraphics[width=\textwidth]{./Figures/lesions.png}
\end{center}

## Current pipeline

Stone et al. _Supervised learning technique for the automated identification of white 
matter hyperintensities in traumatic brain injury_, March, 2016.

\begin{center}
\includegraphics[width=\textwidth]{../Figures/wmhPipeline.png}
\end{center}

## Features

\tiny
\begin{center}
\input{../featureImagesTable}
\end{center}
\normalsize

## Feature images

\begin{center}
$label \sim_{RF} feature_{1} + \ldots + feature_{n}$
\end{center}

<BR><BR>

\begin{center}
\includegraphics[width=\textwidth]{../Figures/featureImages.png}
\end{center}

## Feature importance

\begin{center}
\includegraphics[width=0.6\textwidth]{../Figures/averageLeaveOneOut.png}
\end{center}

## Sample results:  Site 1

\begin{center}
\includegraphics[width=0.75\textwidth]{../Figures/sampleResults.png}
\end{center}

## Leave-one-out evaluation

\begin{center}
\includegraphics[width=\textwidth]{../Figures/evaluationResults.png}
\end{center}

<!--
Evaluation measures for both Stages of the leave-one-out protocol of the described protocol 
in the Methods section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, 
and (d) relative volume difference.   These quantitative assessments are given for three 
quantile ranges spanning the range of the manually-derived lesion volumes.    Overall 
improvement in all three whole lesion-based measuers is seen as the second Stage RF model is 
applied for all three quantile ranges.  The relative volume difference corresponding to the 
Stage 2 results tend to predict a decreased predicted volume over the Stage 1 results.
-->

## Problems with current approach

* Takes a lot of time for new data
    1. Create initial set of feature images
    2. Run it through Stage 1 model
    3. Create more feature images
    4. Run subject through Stage 2 model
* Relatively small data set
* Evaluation/modeling limited to a single site
* Creativity
    * Feature image selection
    * Are we choosing discriminative features?
* Imbalanced data
* No comparison with clinical data

## What about deep learning?

* Potentially quicker for new data
    * Build initial model takes time 
    * No feature images to create for new data
    * GPU
* Optimization learns the features
* Mature packages
    * TensorFlow, CNTK, Torch, Theano, Sci-TK, Caffe, mxnet, Keras
* Need lots of training data

## Current work

1. Rebuild RF models from Site 1 balanced data $\leftarrow$
2. Employ all CENC data $\leftarrow$
3. Apply RF approach $\leftarrow$
4. Check clinical correlations
5. Manually refine data from 2.
6. Re-check clinical correlations
7. Use data from 5. to train deep learning model
8. Re-check clinical correlations

## Imbalanced data

* 7 tissue voxel labels
    * CSF $(n = 208080, 18\%)$
    * gray matter $(n = 457437, 39\%)$
    * white matter $(n = 317788, 27\%)$
    * deep gray matter $(n = 33042, 2.8\%)$
    * brain stem $(n = 17975, 1.5\%)$
    * cerebellum $(n = 1369656, 12\%)$
    * WMH $(n = 901, 0.07\%)$
* SMOTe (Synthetic Minority Over-sampling Technique)
    * For "rare" events:  $\leq 15$\%
    * use bootstrapping and $k$-nearest neighbor